---
title: "Editing state: nudge, snap, and undo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Editing state: nudge, snap, and undo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A `dragmapr_state` is an editable composition object: region and label
offsets can be adjusted from R as well as in the browser. This vignette
covers the small scripted-editing helpers -- nudging one offset, snapping a
whole layout to a grid, validating a state, and keeping an undo/redo history
while you work.

```{r}
library(dragmapr)

state <- d_state(
  region_offsets = data.frame(
    region = c("A", "B"),
    dx_m = c(1234, -567),
    dy_m = c(89, 42)
  )
)
```

## Nudge one offset

`d_nudge()` adds a delta on top of the current offset, like dragging a
region or label slightly in the browser. It returns an updated state with a
bumped version; the input state is unchanged.

```{r}
nudged <- d_nudge(state, "A", dx_m = 500, dy_m = -100)
nudged$region_offsets

label_state <- d_state(
  label_offsets = data.frame(
    label_id = "lbl-a", region = "A", dx_m = 10, dy_m = 20
  )
)
d_nudge(label_state, "lbl-a", dx_m = -5, target = "label")$label_offsets
```

## Snap a layout to a grid

`d_snap_offsets()` rounds every region and label offset to the nearest
multiple of `grid`, tidying a hand-dragged layout into even increments.

```{r}
snapped <- d_snap_offsets(state, grid = 1000)
snapped$region_offsets
```

## Validate a state

`d_validate_state()` returns `TRUE` for a valid state and throws an
informative error otherwise, so it works directly in assertions.

```{r}
d_validate_state(snapped)
```

## Undo and redo

`d_history()` creates an undo/redo stack. Push a checkpoint after each
meaningful edit; `d_undo()` steps back and `d_redo()` steps forward. Pushing
a new checkpoint clears the redo stack, like a text editor.

```{r}
h <- d_history()
d_history_push(h, state)
d_history_push(h, nudged)
d_history_push(h, snapped)

back <- d_undo(h)  # nudged offsets again
back$region_offsets$dx_m

fwd <- d_redo(h)   # snapped offsets again
fwd$region_offsets$dx_m

# Branching clears the redo stack.
branched <- d_nudge(back, "B", dx_m = 2000)
d_history_push(h, branched)
```

The history object has reference semantics, so the push/undo/redo calls
modify it in place -- reassignment is optional.
