---
title: "Example gallery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example gallery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This gallery exists for two reasons: to show the range of draggable plots that
`dragmapr` can create, and to stress the package with different geometry shapes
before real projects depend on it.

## Explodemap-style HHS Fixture

`example_hhs_layout()` bundles the small pieces needed from the explodemap paper
workflow: HHS membership, region names, region colors, and the published
display-offset CSV. The geometry itself is a lightweight projected fixture.

```{r}
library(dragmapr)

hhs <- example_hhs_layout()
render_dragged_map(
  hhs$states,
  region_offsets = hhs$region_offsets,
  region_col = "hhs_region",
  labels = hhs$labels,
  label_offsets = hhs$label_offsets,
  region_palette = hhs$region_colors,
  region_labels = hhs$region_names,
  title = "US Map by HHS Regions"
)
```

## Non-map Panels

The same offset machinery works on any projected `sf` geometry. This panel
example treats rectangles as diagram or dashboard cards.

```{r}
panels <- example_panel_layout()
render_dragged_map(
  panels$panels,
  region_offsets = panels$region_offsets,
  region_col = "group",
  labels = panels$labels,
  label_offsets = panels$label_offsets,
  region_palette = panels$region_colors,
  region_labels = panels$region_names,
  title = "Draggable non-map panels"
)
```

## CSV Round Trip

The key reproducibility test is that offsets can leave the browser as CSV and
come back into R without hidden state.

```{r}
tmp_regions <- tempfile(fileext = ".csv")
tmp_labels <- tempfile(fileext = ".csv")
utils::write.csv(hhs$region_offsets, tmp_regions, row.names = FALSE)
utils::write.csv(hhs$label_offsets, tmp_labels, row.names = FALSE)

render_dragged_map(
  hhs$states,
  region_offsets = tmp_regions,
  region_col = "hhs_region",
  labels = hhs$labels,
  label_offsets = tmp_labels,
  region_palette = hhs$region_colors,
  region_labels = hhs$region_names,
  title = "CSV round trip"
)
```

## Labels, Boxes, And Connectors

This example exercises the newer annotation surface: an info box, a squiggle
connector, thicker connector lines, and plot padding for static export.

```{r}
note <- as_drag_annotations(data.frame(
  label_id = "hhs-9-note",
  region = "9",
  label = "A longer callout can be dragged separately from the region.",
  x = hhs$labels$x[hhs$labels$region == "9"],
  y = hhs$labels$y[hhs$labels$region == "9"]
), width_px = 210, height_px = 88, connector = TRUE, connector_type = "squiggle")

render_dragged_map(
  hhs$states,
  region_offsets = hhs$region_offsets,
  region_col = "hhs_region",
  labels = note,
  label_offsets = data.frame(
    label_id = "hhs-9-note",
    region = "9",
    dx_m = 120000,
    dy_m = 80000
  ),
  region_palette = hhs$region_colors,
  region_labels = hhs$region_names,
  connector_linewidth = 1,
  label_padding = 0.12,
  title = "Annotation box and connector"
)
```

## Shiny

Several Shiny examples are bundled but not evaluated in this vignette because
they launch interactive apps:

```{r, eval = FALSE}
shiny::runApp(system.file("examples", "shiny_draggable_plot.R", package = "dragmapr"))
shiny::runApp(system.file("examples", "shiny_custom_labels.R", package = "dragmapr"))
shiny::runApp(system.file("examples", "shiny_draggable_export.R", package = "dragmapr"))
shiny::runApp(system.file("examples", "shiny_spatial_studio.R", package = "dragmapr"))
shiny::runApp(system.file("examples", "shiny_static_export.R", package = "dragmapr"))
```

`shiny_draggable_export.R` demonstrates a common app pattern: users drag the
interactive plot inside Shiny, the helper sends region and label state back to
the parent app, and the app can export a PNG for a report or document. It also
shows toggles for labels, legends, marker shapes (circle, rounded box, text
only), info-box dimensions, connector style, and connector thickness.

`shiny_spatial_studio.R` is the general spatial workspace. It accepts local
polygon uploads or a URL, including zipped shapefiles, GeoJSON, and GeoPackage
files, then lets users choose grouping/label columns, colors, annotation style
(short labels or info boxes), text size, connector style, legend visibility, and
export formats. Region groups are sorted in natural numeric order throughout the
sidebar, the legend, and the exported CSVs.
