---
title: "Example gallery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example gallery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This gallery exists for two reasons: to show what `dragmapr` can do, and to
stress the package with different geometry shapes before real projects depend
on it.

## Explodemap-style HHS Fixture

`example_hhs_layout()` bundles the small pieces needed from the explodemap paper
workflow: HHS membership, region names, region colors, and the published
display-offset CSV. The geometry itself is a lightweight projected fixture.

```{r}
library(dragmapr)

hhs <- example_hhs_layout()
render_dragged_map(
  hhs$states,
  region_offsets = hhs$region_offsets,
  region_col = "hhs_region",
  labels = hhs$labels,
  label_offsets = hhs$label_offsets,
  region_palette = hhs$region_colors,
  region_labels = hhs$region_names,
  title = "US Map by HHS Regions"
)
```

## Non-map Panels

The same offset machinery works on any projected `sf` geometry. This panel
example treats rectangles as diagram or dashboard cards.

```{r}
panels <- example_panel_layout()
render_dragged_map(
  panels$panels,
  region_offsets = panels$region_offsets,
  region_col = "group",
  labels = panels$labels,
  label_offsets = panels$label_offsets,
  region_palette = panels$region_colors,
  region_labels = panels$region_names,
  title = "Draggable non-map panels"
)
```

## CSV Round Trip

The key reproducibility test is that offsets can leave the browser as CSV and
come back into R without hidden state.

```{r}
tmp_regions <- tempfile(fileext = ".csv")
tmp_labels <- tempfile(fileext = ".csv")
utils::write.csv(hhs$region_offsets, tmp_regions, row.names = FALSE)
utils::write.csv(hhs$label_offsets, tmp_labels, row.names = FALSE)

render_dragged_map(
  hhs$states,
  region_offsets = tmp_regions,
  region_col = "hhs_region",
  labels = hhs$labels,
  label_offsets = tmp_labels,
  region_palette = hhs$region_colors,
  region_labels = hhs$region_names,
  title = "CSV round trip"
)
```
