---
title: "Getting started with dragmapr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with dragmapr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`dragmapr` is built around a deliberately small loop for draggable plots:

1. Start with projected `sf` geometry and a grouping column.
2. Open the draggable browser helper with `drag_map_prototype(open = interactive())`.
3. Drag regions, labels, or annotation boxes until the layout looks right.
4. Copy or download the offset CSVs from the helper panel.
5. Optionally reproduce the layout as a static `ggplot2` image with `render_dragged_map()`.

The interactive helper is the primary deliverable.
Step 5 is optional — use it when you need a reproducible image for a report or publication.

## Toy geometry

All examples below use a projected toy dataset so the workflow can be tested
without downloading shapefiles.

```{r setup}
library(dragmapr)

make_square <- function(x0, y0, size = 100000) {
  sf::st_polygon(list(rbind(
    c(x0, y0), c(x0 + size, y0), c(x0 + size, y0 + size),
    c(x0, y0 + size), c(x0, y0)
  )))
}

regions <- sf::st_sf(
  region = c("North", "South", "East", "West"),
  label  = c("North", "South", "East", "West"),
  geometry = sf::st_sfc(
    make_square(0,       140000),
    make_square(0,            0),
    make_square( 140000,  70000),
    make_square(-140000,  70000),
    crs = 3857
  )
)
```

---

## `drag_map_prototype()` — full parameter reference

### Data

These three arguments identify the geometry and what to group or label.

| Argument | Default | Description |
|---|---|---|
| `x` | *(required)* | Projected `sf` object. Must not be in a geographic (lon/lat) CRS. |
| `region_col` | *(required)* | Column name that defines draggable groups. |
| `label_col` | `region_col` | Column used for label text. Defaults to the grouping column. |

```{r, eval = FALSE}
# Minimal call — one grouping column used for both groups and labels
drag_map_prototype(regions, region_col = "region", open = interactive())

# Separate column for label text
drag_map_prototype(regions, region_col = "region", label_col = "label", open = interactive())
```

---

### Labels

Control whether labels appear and whether they can be dragged independently.

| Argument | Default | Description |
|---|---|---|
| `labels` | `TRUE` | `TRUE` derives one label per region automatically. `FALSE` shows no labels. Supply a data frame from `as_drag_labels()` or `as_drag_annotations()` for custom labels. |
| `draggable_labels` | `TRUE` | When `TRUE`, labels can be moved independently. When `FALSE`, labels follow their region but cannot be repositioned on their own. |

```{r, eval = FALSE}
# No labels
drag_map_prototype(regions, region_col = "region", labels = FALSE, open = interactive())

# Labels follow regions but cannot be dragged separately
drag_map_prototype(regions, region_col = "region",
                   draggable_labels = FALSE, open = interactive())

# Custom label data frame
custom_labels <- make_region_labels(regions, region_col = "region", label_col = "label")
drag_map_prototype(regions, region_col = "region",
                   labels = custom_labels, open = interactive())
```

---

### Label styling

Choose between three marker shapes for short text labels — rounded box, circle,
or text-only — and control their sizes and font.

| Argument | Default | Description |
|---|---|---|
| `label_marker` | `TRUE` | Draw a visible marker shape behind label text. `FALSE` is a shortcut for `label_marker_shape = "none"`. |
| `label_marker_shape` | `"rect"` | Shape behind the label text: `"rect"` (rounded box), `"circle"`, or `"none"` (text only with an invisible drag target). |
| `label_text_size` | `11` | Label text size in browser pixels. Applies to all label types. |
| `label_width` | `64` | Width of rounded-box label markers in pixels. |
| `label_height` | `30` | Height of rounded-box label markers in pixels. |
| `label_radius` | `12` | Circle radius in pixels. Used when `label_marker_shape = "circle"`. |
| `label_box_width` | `150` | Width of draggable annotation box (info-box style) in pixels. |
| `label_box_height` | `72` | Height of draggable annotation box in pixels. |

```{r, eval = FALSE}
# Rounded-box markers (default) — adjust size and text
drag_map_prototype(
  regions, region_col = "region",
  label_marker_shape = "rect",
  label_text_size    = 12,
  label_width        = 80,
  label_height       = 32,
  open = interactive()
)

# Circle markers
drag_map_prototype(
  regions, region_col = "region",
  label_marker_shape = "circle",
  label_radius       = 18,
  label_text_size    = 11,
  open = interactive()
)

# Text only — no visible marker, just draggable text
drag_map_prototype(
  regions, region_col = "region",
  label_marker_shape = "none",
  label_text_size    = 13,
  open = interactive()
)

# Annotation boxes (info-box style with connectors)
notes <- as_drag_annotations(
  make_region_labels(regions, region_col = "region"),
  width_px       = 160,
  height_px      = 70,
  connector      = TRUE,
  connector_type = "elbow"
)
drag_map_prototype(
  regions, region_col = "region",
  labels           = notes,
  label_box_width  = 160,
  label_box_height = 70,
  open = interactive()
)
```

---

### Connectors

Connector lines link a label or annotation box back to its region anchor.
They are only drawn for labels that have `connector = TRUE` in their label data
frame — set that flag via `as_drag_annotations(..., connector = TRUE)` or
`as_drag_labels()`. The line style (`"straight"`, `"elbow"`, `"curve"`,
`"squiggle"`) is stored per-label in the `connector_type` column.

| Argument | Default | Description |
|---|---|---|
| `connector_color` | `"#334155"` | Browser connector line colour. |
| `connector_linewidth` | `1.3` | Connector line thickness in browser pixels. |
| `connector_linetype` | `"solid"` | Browser connector line pattern: `"solid"`, `"dashed"`, or `"dotted"`. |
| `connector_endpoint` | `"none"` | Browser connector endpoint: `"none"` or `"arrow"`. |
| `connector_smart` | `FALSE` | Let the browser choose straight, elbow, or curved connector geometry from the current label displacement. |

```{r, eval = FALSE}
drag_map_prototype(
  regions, region_col = "region",
  labels              = notes,   # labels with connector = TRUE
  connector_color     = "#334155",
  connector_linewidth = 2,
  connector_linetype  = "dashed",
  connector_endpoint  = "arrow",
  connector_smart     = TRUE,
  open = interactive()
)
```

---

### Movement context

Movement context is optional and off by default. It helps explain where a moved
region came from without changing the underlying offset data.

| Argument | Default | Description |
|---|---|---|
| `show_origin_outlines` | `FALSE` | Draw the original outline for regions with non-zero offsets. |
| `show_movement_connectors` | `FALSE` | Draw a line from each moved region's original representative point to its current location. |
| `movement_connector_color` | `"#64748b"` | Movement connector colour. |
| `movement_connector_opacity` | `0.72` | Movement connector opacity from 0 to 1. |
| `movement_connector_linewidth` | `1.4` | Browser movement connector thickness in pixels. |
| `movement_connector_linetype` | `"solid"` | Movement connector pattern: `"solid"`, `"dashed"`, or `"dotted"`. |
| `movement_connector_endpoint` | `"closed"` | Movement connector endpoint: `"closed"`, `"open"`, or `"none"`. |
| `show_drag_trail` | `FALSE` | Browser-only fading outline trail while a region is actively dragged. |

```{r, eval = FALSE}
drag_map_prototype(
  regions, region_col = "region",
  show_origin_outlines = TRUE,
  show_movement_connectors = TRUE,
  movement_connector_color = "#64748b",
  movement_connector_linetype = "dashed",
  movement_connector_endpoint = "open",
  show_drag_trail = TRUE,
  open = interactive()
)
```

Static exports support origin outlines and movement connectors too:

```{r, eval = FALSE}
render_dragged_map(
  regions,
  region_offsets = region_offsets,
  region_col = "region",
  show_origin_outlines = TRUE,
  show_movement_connectors = TRUE,
  movement_connector_linetype = "dotted",
  movement_connector_endpoint = "closed"
)
```

---

### Initial positions

Supply previously exported offset tables to pre-load a saved layout so
dragging can continue from where it left off.

| Argument | Default | Description |
|---|---|---|
| `region_offsets` | `NULL` | Data frame with columns `region`, `dx_m`, `dy_m`. All regions start at zero when `NULL`. |
| `label_offsets` | `NULL` | Data frame with columns `label_id`, `region`, `dx_m`, `dy_m`. All labels start at zero when `NULL`. |

```{r, eval = FALSE}
# Resume from previously downloaded CSVs
region_csv <- read_offsets("drag_region_offsets.csv")
label_csv  <- read_label_state("drag_label_offsets.csv")

drag_map_prototype(
  regions, region_col = "region",
  region_offsets = region_csv,
  label_offsets  = label_csv,
  open = interactive()
)
```

`read_label_offsets()` is retained as an alias for `read_label_state()`, and
`apply_label_offsets()` is retained as an alias for `apply_label_state()`.
Likewise, `make_labels()` remains an alias for `make_region_labels()` for older
scripts.

If the offsets were exported as a Spatial Studio project ZIP, use the one-line
project renderer instead of wiring the CSV files together manually:

```{r, eval = FALSE}
render_dragmapr_project(
  "dragmapr-project.zip",
  file = "dragged-map.png"
)
```

---

### RStudio addin

In RStudio, install the package and use **Addins > Launch dragmapr** to open a
compact prototype gadget. The addin looks for `sf` objects in `.GlobalEnv`,
lets you choose the region and label columns, and assigns `region_offsets` and
`label_offsets` when you click **Done**.
Create or load the `sf` object before opening the addin. If you call
`dragmapr_addin()` directly, pass `env =` to work with another environment.

---

### Colour palette

| Argument | Default | Description |
|---|---|---|
| `region_palette` | `NULL` | Named character vector mapping each region value to a hex colour string. Uses a built-in 8-colour cycling palette when `NULL`. |

Matching the interactive palette to a `render_dragged_map()` call makes the
draggable preview and the final static image look identical.

```{r, eval = FALSE}
my_palette <- c(
  North = "#4C78A8",
  South = "#F58518",
  East  = "#54A24B",
  West  = "#B279A2"
)

drag_map_prototype(
  regions, region_col = "region",
  region_palette = my_palette,
  show_legend    = TRUE,
  open = interactive()
)
```

---

### Legend

A compact colour legend can be embedded directly inside the browser helper.
It is hidden by default because many maps are labelled and need no legend.

| Argument | Default | Description |
|---|---|---|
| `show_legend` | `FALSE` | Show a colour legend in the helper. |
| `max_legend_keys` | `25` | Suppress the legend automatically when the number of groups exceeds this value. Set to `Inf` to always show it regardless of group count. |
| `legend_position` | `"bottom"` | Where the legend appears: `"bottom"`, `"top"`, `"left"`, `"right"`, or `"none"`. |
| `legend_title` | `"Region"` | Title shown above the legend keys. |
| `legend_values` | `NULL` | Optional character vector of region values to include in the legend. `NULL` includes all values. |

```{r, eval = FALSE}
drag_map_prototype(
  regions, region_col = "region",
  region_palette  = my_palette,
  show_legend     = TRUE,
  legend_position = "bottom",   # "top" | "left" | "right" | "none"
  legend_title    = "Region",
  legend_values   = c("North", "East"),
  max_legend_keys = 25,
  open = interactive()
)
```

Use `label_values` when an app should temporarily display only selected labels
while preserving all label offsets:

```{r, eval = FALSE}
drag_map_prototype(
  regions, region_col = "region",
  label_values = c("North", "East"),
  open = interactive()
)
```

---

### Output file and display

| Argument | Default | Description |
|---|---|---|
| `file` | `"drag-map.html"` | File path for the written standalone HTML. |
| `open` | `FALSE` | Open the file in the default browser via `browseURL()` after writing. |
| `side_panel` | `TRUE` | Show the built-in copy/download side panel. Shiny apps that provide their own export controls can set this to `FALSE` to reclaim the full canvas. |
| `map_background` | `"white"` | Helper background: `"white"`, `"transparent"`, `"light_grid"`, or `"dark"`. |

```{r, eval = FALSE}
# Write to a chosen path without opening (e.g. CI, server, or Shiny)
drag_map_prototype(
  regions, region_col = "region",
  file       = "output/my_map.html",
  map_background = "light_grid",
  side_panel = TRUE,
  open       = FALSE
)

# Embedded in a Shiny app — hide the panel; the app provides its own controls
drag_map_prototype(
  regions, region_col = "region",
  side_panel = FALSE,
  file       = tempfile(fileext = ".html")
)
```

---

## Optional: Static rendering

When you need a reproducible `ggplot2` image, pass the exported offset tables
to `render_dragged_map()`. The data frames below stand in for downloaded CSVs:

```{r}
region_offsets <- data.frame(
  region = c("North", "South", "East", "West"),
  dx_m   = c(0, 0,  60000, -60000),
  dy_m   = c(50000, -50000, 0, 0)
)

label_state <- data.frame(
  label_id = c("North", "South", "East", "West"),
  region   = c("North", "South", "East", "West"),
  dx_m     = c(0, 0,  25000, -25000),
  dy_m     = c(30000, -30000, 0, 0)
)

render_dragged_map(
  regions,
  region_offsets = region_offsets,
  region_col     = "region",
  label_col      = "label",
  label_offsets  = label_state,
  title          = "Synthetic draggable map"
)
```

The same annotation types available in the interactive helper work in static output:

```{r}
notes <- as_drag_annotations(data.frame(
  label_id = "east-note",
  region   = "East",
  label    = "The East label can become a longer draggable note.",
  x = 210000, y = 120000
), connector = TRUE, connector_type = "squiggle")

render_dragged_map(
  regions,
  region_offsets      = region_offsets,
  region_col          = "region",
  labels              = notes,
  label_offsets       = data.frame(
    label_id = "east-note", region = "East",
    dx_m = 80000, dy_m = 45000
  ),
  connector_linewidth = 0.8,
  connector_linetype  = "dotted",
  connector_endpoint  = "arrow",
  legend_title        = "Region",
  map_background      = "light_grid",
  title               = "Static output with an annotation box"
)
```

---

## Edit the source features

Sometimes the right display map is not just a moved map. It is a lighter map:
one or more features are removed, or a corrected feature is swapped in before
the final layout is composed.

```{r, eval = FALSE}
feature_rows <- spatial_feature_table(regions, key_col = "region")
feature_rows[, c(".feature_id", ".row", ".geometry_type")]

regions_without_north <- remove_spatial_features(
  regions,
  ids = "North",
  key_col = "region"
)

regions_only_south_west <- keep_spatial_features(
  regions,
  ids = c("South", "West"),
  key_col = "region"
)
```

The editing helpers return ordinary `sf` objects, so the result can go straight
back into the draggable helper:

```{r, eval = FALSE}
drag_map_prototype(
  regions_without_north,
  region_col = "region",
  open = interactive()
)
```

## A small Shiny editor

This is the same loop inside a minimal Shiny app: render the widget, receive
state from the browser, and push a live display update from the server.

```{r, eval = FALSE}
library(shiny)
library(dragmapr)

ui <- fluidPage(
  checkboxInput("origin", "Show origin outlines", TRUE),
  dragmaprOutput("map", height = "650px"),
  verbatimTextOutput("state")
)

server <- function(input, output, session) {
  current_state <- reactiveVal(dragmapr_state(crs = 3857))

  output$map <- renderDragmapr({
    dragmapr_widget(
      regions,
      region_col = "region",
      state = current_state(),
      display_options = dragmapr_display_options(
        show_origin_outlines = isTRUE(input$origin)
      )
    )
  })

  observeEvent(input$map_state, {
    current_state(dragmapr_widget_state(input$map_state))
  })

  observeEvent(input$origin, {
    updateDragmapr(session, "map", show_origin_outlines = input$origin)
  })

  output$state <- renderPrint(summary(current_state()))
}

shinyApp(ui, server)
```

## Working with your own data

Use projected polygon data for editing. If your source file is longitude and
latitude, run it through `prepare_dragmapr_sf()` first so offsets are measured
in metres:

```{r, eval = FALSE}
my_sf <- read_dragmapr_sf_url("https://example.com/regions.geojson")
my_sf <- prepare_dragmapr_sf(my_sf)

drag_map_prototype(
  my_sf,
  region_col = "region_name",
  label_col = "short_name",
  map_background = "light_grid",
  open = interactive()
)
```

The package keeps the display edits separate from the source geometry. That
separation is the useful part: you can drag, remove, restore, export, and render
without hiding how the final map was made.
