---
title: "Getting started with dragmapr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with dragmapr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`dragmapr` is built around a deliberately small loop for draggable plots:

1. Start with projected `sf` geometry and a grouping column.
2. Open a draggable browser helper with `drag_map_prototype(open = TRUE)`.
3. Drag regions, labels, or annotation boxes until the layout works.
4. Use the draggable plot directly, or copy/download the offset CSVs.
5. Optionally reproduce the layout as a static `ggplot2` image from those CSVs.

The interactive browser helper is the primary deliverable. Step 5 is optional -
use it when you need a reproducible image for a report or publication.

## A Synthetic Map

The package examples use projected toy geometries so the workflow can be tested
without downloading shapefiles.

```{r}
library(dragmapr)

make_square <- function(x0, y0, size = 100000) {
  sf::st_polygon(list(rbind(
    c(x0, y0),
    c(x0 + size, y0),
    c(x0 + size, y0 + size),
    c(x0, y0 + size),
    c(x0, y0)
  )))
}

regions <- sf::st_sf(
  region = c("North", "South", "East", "West"),
  label = c("North", "South", "East", "West"),
  geometry = sf::st_sfc(
    make_square(0, 140000),
    make_square(0, 0),
    make_square(140000, 70000),
    make_square(-140000, 70000),
    crs = 3857
  )
)
```

Open the interactive helper in the browser - this is the main step:

```{r, eval = FALSE}
drag_map_prototype(
  regions,
  region_col = "region",
  label_col = "label",
  labels = TRUE,
  draggable_labels = TRUE,
  open = TRUE
)
```

Drag regions and labels until the layout looks right. The helper panel shows
live offset tables. Use **Copy Regions** / **Copy Labels** or the download
buttons to capture the two CSVs. Region offsets move grouped geometries; label
offsets move labels and annotation boxes after their region has moved.

## Optional: Static Rendering

If you need a reproducible `ggplot2` image, supply the exported offsets to
`render_dragged_map()`. The data frames below stand in for the downloaded CSVs:

```{r}
region_offsets <- data.frame(
  region = c("North", "South", "East", "West"),
  dx_m = c(0, 0, 60000, -60000),
  dy_m = c(50000, -50000, 0, 0)
)

label_state <- data.frame(
  label_id = c("North", "South", "East", "West"),
  region = c("North", "South", "East", "West"),
  dx_m = c(0, 0, 25000, -25000),
  dy_m = c(30000, -30000, 0, 0)
)

plot <- render_dragged_map(
  regions,
  region_offsets = region_offsets,
  region_col = "region",
  label_col = "label",
  label_offsets = label_state,
  title = "Synthetic draggable map"
)
plot
```

Static output can include the same label choices as the draggable helper:

```{r}
notes <- as_drag_annotations(data.frame(
  label_id = "east-note",
  region = "East",
  label = "The East label can become a longer draggable note.",
  x = 210000,
  y = 120000
), connector = TRUE, connector_type = "squiggle")

render_dragged_map(
  regions,
  region_offsets = region_offsets,
  region_col = "region",
  labels = notes,
  label_offsets = data.frame(
    label_id = "east-note",
    region = "East",
    dx_m = 80000,
    dy_m = 45000
  ),
  connector_linewidth = 0.8,
  title = "Static output with an annotation box"
)
```

## Running Examples

Installed example scripts live under `system.file("examples", package =
"dragmapr")`. To smoke test the full package surface:

```{r, eval = FALSE}
source(system.file("examples", "smoke_examples.R", package = "dragmapr"))
```

Shiny examples are available separately because they launch interactive apps.
See the [Shiny workflows vignette](shiny-workflows.html) for a walkthrough, or
run them directly:

```{r, eval = FALSE}
shiny::runApp(system.file("examples", "shiny_draggable_plot.R",   package = "dragmapr"))
shiny::runApp(system.file("examples", "shiny_custom_labels.R",    package = "dragmapr"))
shiny::runApp(system.file("examples", "shiny_draggable_export.R", package = "dragmapr"))
shiny::runApp(system.file("examples", "shiny_spatial_studio.R",   package = "dragmapr"))
shiny::runApp(system.file("examples", "shiny_static_export.R",    package = "dragmapr"))
```
