---
title: "Getting started with dragmapr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with dragmapr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`dragmapr` is built around a deliberately small loop:

1. Start with projected `sf` geometry and a grouping column.
2. Write a draggable HTML helper.
3. Drag regions and labels.
4. Save offset CSVs.
5. Render the final static image from the original data plus those CSVs.

This keeps manual layout decisions reviewable in Git. The browser is an editor,
not the source of truth.

## A Synthetic Map

The package examples use projected toy geometries so the workflow can be tested
without downloading shapefiles.

```{r}
library(dragmapr)

make_square <- function(x0, y0, size = 100000) {
  sf::st_polygon(list(rbind(
    c(x0, y0),
    c(x0 + size, y0),
    c(x0 + size, y0 + size),
    c(x0, y0 + size),
    c(x0, y0)
  )))
}

regions <- sf::st_sf(
  region = c("North", "South", "East", "West"),
  label = c("North", "South", "East", "West"),
  geometry = sf::st_sfc(
    make_square(0, 140000),
    make_square(0, 0),
    make_square(140000, 70000),
    make_square(-140000, 70000),
    crs = 3857
  )
)
```

Write the helper HTML:

```{r, eval = FALSE}
drag_map_prototype(
  regions,
  region_col = "region",
  label_col = "label",
  file = "basic_drag_helper.html"
)
```

After dragging, the helper exports a region-offset CSV and a label-offset CSV.
They can be supplied as data frames or file paths:

```{r}
region_offsets <- data.frame(
  region = c("North", "South", "East", "West"),
  dx_m = c(0, 0, 60000, -60000),
  dy_m = c(50000, -50000, 0, 0)
)

label_offsets <- data.frame(
  label_id = c("North", "South", "East", "West"),
  region = c("North", "South", "East", "West"),
  dx_m = c(0, 0, 25000, -25000),
  dy_m = c(30000, -30000, 0, 0)
)

plot <- render_dragged_map(
  regions,
  region_offsets = region_offsets,
  region_col = "region",
  label_col = "label",
  label_offsets = label_offsets,
  title = "Synthetic draggable map"
)
plot
```

## Running Examples

Installed example scripts live under `system.file("examples", package =
"dragmapr")`. To smoke test the full package surface:

```{r, eval = FALSE}
source(system.file("examples", "smoke_examples.R", package = "dragmapr"))
```
