---
title: "HHS placeholder shapes demo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HHS placeholder shapes demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This short video shows the embedded HHS placeholder shapes workflow in Spatial
Studio. It uses the bundled HHS-style fixture as an actual map-shaped starting
point for dragging regions, adjusting labels, and exporting reproducible
artifacts.

```{r hhs-video, echo = FALSE, results = "asis"}
in_pkgdown <- identical(Sys.getenv("IN_PKGDOWN"), "true") ||
  isTRUE(getOption("pkgdown.in_pkgdown"))

if (in_pkgdown) {
  cat('
<div class="dragmapr-demo-video">
  <video controls preload="metadata" poster="../hhs-placeholder-shapes-poster.png" src="../hhs-placeholder-shapes-demo.mp4">
    <p>
      Your browser does not support embedded video.
      <a href="../hhs-placeholder-shapes-demo.mp4">Download the HHS placeholder shapes demo video.</a>
    </p>
  </video>
</div>
')
} else {
  cat("[Watch the HHS placeholder shapes demo on the pkgdown site](https://prigasg.github.io/dragmapr/articles/hhs-placeholder-shapes-demo.html).")
}
```

The same HHS-style fixture is available from R with `example_hhs_layout()`:

```{r, eval = FALSE}
library(dragmapr)

hhs <- example_hhs_layout()

drag_map_prototype(
  hhs$states,
  region_col = "hhs_region",
  labels = hhs$labels,
  region_offsets = hhs$region_offsets,
  label_offsets = hhs$label_offsets,
  region_palette = hhs$region_colors,
  show_legend = TRUE,
  legend_title = "HHS region",
  open = TRUE
)
```

For a full Spatial Studio session, launch the bundled app:

```{r, eval = FALSE}
shiny::runApp(system.file(
  "examples",
  "shiny_spatial_studio.R",
  package = "dragmapr"
))
```
