---
title: "Labels and static output"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Labels and static output}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Labels have their own data model in `dragmapr` because draggable plots often
need two kinds of movement:

- region offsets move all geometries in a group;
- label offsets nudge only the text marker after the region has moved.

If users want a static image after the draggable edit, the same data model can
be rendered back through `ggplot2`. This means a Shiny app is optional: once
the region and label offset CSVs exist, `render_dragged_map()` can reconstruct
the layout later in a script, report, or document workflow.

Labels can also be configured or omitted when the draggable helper is written:

```{r, eval = FALSE}
drag_map_prototype(x, "region", labels = FALSE)
drag_map_prototype(x, "region", label_col = "name", draggable_labels = FALSE)
drag_map_prototype(x, "region", label_col = "name", label_marker = FALSE)
drag_map_prototype(x, "region", label_col = "name", label_radius = 16, label_text_size = 14)
```

## Label Concepts

`dragmapr` separates labels into three related concepts:

- `make_region_labels()` derives one default label anchor per group.
- `as_drag_labels()` accepts user-supplied labels and preserves extra metadata.
- `as_drag_annotations()` creates draggable info boxes for longer notes.
- `read_label_state()` and `apply_label_state()` restore label movements from
  CSV state exported by the draggable helper.

By default, region labels use `sf::st_point_on_surface()` so anchors are likely
to fall inside grouped geometry.

```{r}
library(dragmapr)

hhs <- example_hhs_layout()
head(hhs$labels)
```

User-supplied labels are plain data:

```text
label_id,region,label,x,y
```

They may also carry extra columns for Shiny/D3 use:

```{r}
as_drag_labels(data.frame(
  label_id = "note-1",
  region = "3",
  label = "Custom note",
  x = hhs$labels$x[3],
  y = hhs$labels$y[3],
  tooltip = "Shown by a custom D3/Shiny layer"
))
```

## Info Boxes And Text-only Labels

Info boxes are just label rows with `label_type = "box"` plus browser box
dimensions. `as_drag_annotations()` fills those details for you:

```{r}
note <- as_drag_annotations(data.frame(
  label_id = "region-3-note",
  region = "3",
  label = "Region 3 has a longer note that reads better in a box.",
  x = hhs$labels$x[3],
  y = hhs$labels$y[3]
), width_px = 180, height_px = 84)

note
```

Ordinary labels can be rendered with a circular marker or as text-only labels.
Text-only labels remain draggable in the browser helper because `dragmapr` adds
an invisible drag target behind the text.

```{r}
render_dragged_map(
  hhs$states,
  region_offsets = hhs$region_offsets,
  region_col = "hhs_region",
  labels = hhs$labels,
  label_offsets = hhs$label_offsets,
  region_palette = hhs$region_colors,
  region_labels = hhs$region_names,
  show_label_marker = FALSE,
  title = "Text-only labels"
)
```

## Connector Lines

Labels and annotation boxes can have connector lines. The connector starts from
the original label anchor by default and ends just outside the visible label or
box. Optional `connector_start_x` / `connector_start_y` columns can override the
start point, while `connector_mid_x` / `connector_mid_y` can define a breakpoint
for elbow or curved connectors.

Supported connector styles are `"straight"`, `"elbow"`, `"curve"`, and
`"squiggle"`.

```{r}
note$connector <- TRUE
note$connector_type <- "squiggle"

render_dragged_map(
  hhs$states,
  region_offsets = hhs$region_offsets,
  region_col = "hhs_region",
  labels = note,
  label_offsets = data.frame(
    label_id = "region-3-note",
    region = "3",
    dx_m = 90000,
    dy_m = 70000
  ),
  region_palette = hhs$region_colors,
  region_labels = hhs$region_names,
  connector_linewidth = 0.9,
  label_padding = 0.12,
  title = "Annotation box with connector"
)
```

`label_padding` expands static plot limits around displaced labels and
connectors, which helps prevent exported images from clipping callouts.

The exported label-offset table is also plain data:

```text
label_id,region,dx_m,dy_m
```

## Region Movement Plus Label Movement

Region offsets are applied to both geometry and default label anchors. Label
offsets are applied after that.

```{r}
label_offsets <- hhs$label_offsets
label_offsets$dx_m[label_offsets$label_id == "3"] <- -45000
label_offsets$dy_m[label_offsets$label_id == "3"] <- -30000

render_dragged_map(
  hhs$states,
  region_offsets = hhs$region_offsets,
  region_col = "hhs_region",
  labels = hhs$labels,
  label_offsets = label_offsets,
  region_palette = hhs$region_colors,
  region_labels = hhs$region_names,
  title = "HHS regions with label nudge"
)
```

## Saving Static Images

Supplying `file` passes the finished plot to `ggplot2::ggsave()`.

```{r, eval = FALSE}
render_dragged_map(
  hhs$states,
  region_offsets = "drag_region_offsets.csv",
  region_col = "hhs_region",
  labels = hhs$labels,
  label_offsets = "drag_label_offsets.csv",
  file = "hhs_dragged_layout.png",
  width = 9,
  height = 6
)
```
