---
title: "Labels and static output"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Labels and static output}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Labels have their own data model in `dragmapr` because final map composition
often needs two kinds of movement:

- region offsets move all geometries in a group;
- label offsets nudge only the text marker after the region has moved.

This mirrors the workflow used by the original explodemap paper scripts while
keeping the package independent of explodemap.

## Label Anchors

`make_labels()` derives one label anchor per region. By default it uses
`sf::st_point_on_surface()` so anchors are likely to fall inside the grouped
geometry.

```{r}
library(dragmapr)

hhs <- example_hhs_layout()
head(hhs$labels)
```

The label table is plain data:

```text
label_id,region,label,x,y
```

The exported label-offset table is also plain data:

```text
label_id,region,dx_m,dy_m
```

## Region Movement Plus Label Movement

Region offsets are applied to both geometry and default label anchors. Label
offsets are applied after that.

```{r}
label_offsets <- hhs$label_offsets
label_offsets$dx_m[label_offsets$label_id == "3"] <- -45000
label_offsets$dy_m[label_offsets$label_id == "3"] <- -30000

render_dragged_map(
  hhs$states,
  region_offsets = hhs$region_offsets,
  region_col = "hhs_region",
  labels = hhs$labels,
  label_offsets = label_offsets,
  region_palette = hhs$region_colors,
  region_labels = hhs$region_names,
  title = "HHS regions with label nudge"
)
```

## Saving Static Images

Supplying `file` passes the finished plot to `ggplot2::ggsave()`.

```{r, eval = FALSE}
render_dragged_map(
  hhs$states,
  region_offsets = "drag_region_offsets.csv",
  region_col = "hhs_region",
  labels = hhs$labels,
  label_offsets = "drag_label_offsets.csv",
  file = "hhs_dragged_layout.png",
  width = 9,
  height = 6
)
```
