---
title: "Shiny workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Shiny workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`dragmapr` is useful in Shiny when the draggable plot is the interface itself.
Users can move regions, panels, or labels directly instead of typing numeric
offsets.

## Embed A Draggable Plot

The simplest Shiny pattern is to write a helper HTML file into a Shiny resource
directory and show it in an iframe.

```{r, eval = FALSE}
shiny::runApp(system.file("examples", "shiny_draggable_plot.R", package = "dragmapr"))
```

This is useful when the app only needs to let users explore or compose a layout.

## Custom Labels

Apps can provide their own label table instead of using one label per region.
This is useful for annotations, review notes, callouts, or workflow-specific
metadata. Labels can be ordinary text labels, text-only draggable labels, or
annotation boxes created with `as_drag_annotations()`.

```{r, eval = FALSE}
shiny::runApp(system.file("examples", "shiny_custom_labels.R", package = "dragmapr"))
```

The user-supplied label table is created with `as_drag_labels()` and can carry
extra columns for app-specific behavior. Connector columns such as `connector`,
`connector_type`, `connector_start_x`, and `connector_mid_x` are also preserved,
so apps can let users choose straight, elbow, curved, or squiggle leader lines.

## Export A Report Image

Some apps need a static image after the user has finished dragging. The helper
posts offset state to its Shiny parent window. The Shiny app can then render a
preview and expose a PNG download.

```{r, eval = FALSE}
shiny::runApp(system.file("examples", "shiny_draggable_export.R", package = "dragmapr"))
```

This pattern is useful for report builders, document workflows, or review apps
where the interactive layout is the editing surface and the exported PNG is the
deliverable.

The bundled `shiny_draggable_export.R` app intentionally has many controls
because it acts as a smoke test for the package surface:

- show or hide labels;
- switch between short labels and info boxes;
- show text labels with or without circular markers;
- change info-box width and height;
- show connector lines;
- choose straight, elbow, curve, or squiggle connectors;
- adjust connector thickness;
- show or hide the region legend;
- download a static PNG reconstructed from the current region and label offset
  state.

The important design point is that the Shiny app does not hold magic layout
state. The parent app receives the same region and label offset tables that a
non-Shiny workflow can copy or download from the helper.

## Static Only

If offsets have already been saved, an app can skip the draggable helper and
only provide static preview/export.

```{r, eval = FALSE}
shiny::runApp(system.file("examples", "shiny_static_export.R", package = "dragmapr"))
```

## Spatial Studio

`shiny_spatial_studio.R` is the most complete Shiny example. It is a small
spatial workspace rather than a fixed demo: users upload local polygon data or
enter a URL, pick grouping and label columns, edit colors, drag boundaries,
switch between short labels and info boxes, control text size and connector
style, preview a static render, and download the current artifacts.

```{r, eval = FALSE}
shiny::runApp(system.file("examples", "shiny_spatial_studio.R", package = "dragmapr"))
```

Supported inputs: zipped shapefiles, shapefile sidecar files (`.shp` + `.dbf` +
`.shx`), GeoJSON, and GeoPackage — either uploaded locally or downloaded from
a URL. The app reads geometry with `sf::st_read()`, repairs it with
`sf::st_make_valid()`, and transforms longitude/latitude data to EPSG:3857 so
metre offsets work correctly. Region group names are sorted using a natural
(numeric-aware) order, so "1", "2", ..., "10" are displayed in the right
sequence rather than lexicographic order.

The label sidebar consolidates controls so only sliders relevant to the current
label type are shown: text size is always visible; marker size (width + height)
appears only for rounded-box labels; circle radius only for circle labels; box
dimensions only for info boxes.

Available downloads from the studio sidebar:

- **PNG** — static render from `render_dragged_map()`
- **Region CSV** — current region offsets, importable by any later session
- **Label CSV** — current label offsets
- **GeoJSON / GPKG** — the adjusted sf geometry with offsets applied
- **HTML helper** — the standalone D3 drag file you can share or reopen
- **Bundle ZIP** — all of the above plus `palette.csv` and `metadata.json`;
  enough to fully reconstruct the layout in a new session or hand off to a
  collaborator

The app names its central reactive values in the State tab so they are easy to
find when adapting the code:

- `source_sf()` — raw sf from upload / URL / demo
- `projected_sf()` — after `prepare_dragmapr_sf()`
- `region_col()`, `label_col()` — chosen columns
- `label_table()` — styled label data frame passed to `drag_map_prototype()`
- `region_state()`, `label_state()` — current drag offsets as data frames
- `region_palette()` — named colour vector
- `current_plot()` — `ggplot2` object ready to save

### Loading veil

The studio shows a loading veil while data is being read and the D3 helper is
being built. The veil is dismissed when the helper iframe signals it has
finished its first `render()` call by posting a `dragmapr-ready` message to the
parent page. This avoids the race condition where the browser fires the iframe
`load` event before the parent's listener has been attached.

## Reusable Shiny Helpers

The spatial studio internally uses four package-level helpers that are also
available directly for custom apps.

**`read_dragmapr_sf_upload(upload)`** wraps a [shiny::fileInput()] result -
including multi-file shapefile sidecar uploads and zip archives - into a single
`sf::st_read()` call.  Returns `NULL` when the upload is empty so callers can
fall back to demo data.

**`read_dragmapr_sf_url(url, timeout = 60)`** downloads a spatial file from a
URL into a temporary directory, unpacks zip archives, and returns an `sf`
object.  Raises a descriptive error on network or format failures.

**`prepare_dragmapr_sf(x, target_crs = 3857)`** repairs invalid geometry, keeps
only polygon types, assigns a fallback CRS when none is present, and reprojects
geographic data to the target projected CRS so metre offsets are meaningful.

**`dragmapr_iframe_bridge(...)`** returns a JavaScript string that installs the
`postMessage` listener and polling loop needed to relay drag state from the
helper iframe back to Shiny inputs.  Wrap it in `tags$head(tags$script(HTML(
dragmapr_iframe_bridge())))` in your UI.  The function accepts `region_input`,
`label_input`, `slow_poll_ms`, and `fast_poll_ms` arguments to customise the
input names and timing.

```{r, eval = FALSE}
library(shiny)
library(dragmapr)

ui <- fluidPage(
  tags$head(tags$script(HTML(dragmapr_iframe_bridge()))),
  uiOutput("helper")
)

server <- function(input, output, session) {
  helper_dir <- tempfile("myapp_")
  dir.create(helper_dir)
  shiny::addResourcePath("myapp_static", helper_dir)

  x <- prepare_dragmapr_sf(my_sf)
  drag_map_prototype(x, region_col = "region",
                     file = file.path(helper_dir, "helper.html"))

  output$helper <- renderUI(
    tags$iframe(src = "myapp_static/helper.html",
                style = "width:100%;height:700px;border:none;")
  )
}

shinyApp(ui, server)
```
