---
title: "State-first workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{State-first workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

The recommended way to use dragmapr is to treat a layout as an editable
**composition object** -- a `dragmapr_state` -- and carry it through three
steps:

```
compute  ->  compose  ->  render
(layout)     (edit)       (static / interactive)
```

The state is plain, reproducible data: region and label offsets (as metre
deltas), the projected `crs`, a `geometry_id`, the `selected_feature`, and a
monotonically increasing `version`. Because the geometry and the editorial
state are kept separate, you can recompute the layout while preserving manual
edits, and the *same* state object drives the interactive editor, the static
renderer, and any persistence.

The older route -- exporting offset CSVs and calling `as_dragmapr()` -- still
works, but it is now the low-level escape hatch rather than the main road.

## 1. Compute

In a full pipeline the layout comes from explodemap, which computes a
mathematically valid exploded layout and hands it over as a state:

```{r compute-explodemap, eval = FALSE}
library(explodemap)

layout <- explode_grouped(my_sf, region_col = "region")
state  <- as_dragmapr_state(layout)   # dx_m/dy_m deltas + crs + geometry_id
```

For a self-contained example we build the equivalent state directly, so this
vignette runs with only dragmapr installed:

```{r compute-self-contained}
library(dragmapr)

make_square <- function(x0, y0, size = 100000) {
  sf::st_polygon(list(rbind(
    c(x0, y0), c(x0 + size, y0), c(x0 + size, y0 + size),
    c(x0, y0 + size), c(x0, y0)
  )))
}

regions <- sf::st_sf(
  region = c("North", "South", "East", "West"),
  geometry = sf::st_sfc(
    make_square(0, 140000), make_square(0, 0),
    make_square(140000, 70000), make_square(-140000, 70000),
    crs = 3857
  )
)

state <- dragmapr_state(
  region_offsets = data.frame(
    region = c("North", "South", "East", "West"),
    dx_m   = c(0, 0, 60000, -60000),
    dy_m   = c(50000, -50000, 0, 0)
  ),
  crs = 3857,
  geometry_id = "synthetic-quad-v1"
)
state
```

## 2. Compose

`dragmapr_edit()` is the friendly front door to the interactive editor. It
accepts a projected `sf`, an explodemap `grouped_exploded_map`, or a
`dragmapr_layout`, seeded with the state:

```{r compose, eval = FALSE}
dragmapr_edit(regions, region_col = "region", state = state)
```

Inside Shiny, the widget reports every edit through an input named
`paste0(outputId, "_state")`. Rebuild a `dragmapr_state` from it with
`dragmapr_widget_state()` so the server always holds the live composition:

```{r compose-shiny, eval = FALSE}
output$map <- renderDragmapr(dragmapr_edit(regions, "region", state = state))

observeEvent(input$map_state, {
  state <- dragmapr_widget_state(input$map_state)
})

# Push a selection from R back into the browser:
updateDragmapr(session, "map", selected_feature = "East")
```

## 3. Render

Because the state is just data, you can persist it and reproduce the layout as a
static `ggplot2` image at any time -- no browser session and no recomputation:

```{r render}
path <- tempfile(fileext = ".json")
write_dragmapr_state(state, path)
state2 <- read_dragmapr_state(path)

render_dragged_map(
  regions,
  region_col = "region",
  state = state2,
  title = "Composed layout (state-first)"
)
```

## The state object

A `dragmapr_state` carries:

- `region_offsets` / `label_offsets` -- metre deltas (`dx_m`, `dy_m`) keyed by a
  stable join key;
- `crs` -- the projected CRS the deltas are expressed in (EPSG or WKT), so a
  saved state is safe to reapply later;
- `geometry_id` -- a provenance tag for the geometry the state was composed
  against;
- `selected_feature`, `expanded_groups`, `view` -- editor/dashboard state;
- `version` -- a revision that bumps monotonically as edits accumulate.

The full lifecycle (`validate_`, `merge_`, `snapshot_`/`restore_`,
`inherit_drag_offsets()`, `collapse_drag_offsets()`, and JSON `read_`/`write_`)
operates on this object. For a complete runnable round-trip, see the Shiny app
in `system.file("examples/full_state_roundtrip.R", package = "dragmapr")`.

For a fuller cross-package example that starts in `explodemap`, uses
layout-quality diagnostics and label-aware parameter search, then composes and
renders with `dragmapr`, run:

```{r full-cross-package, eval = FALSE}
source(system.file(
  "examples/explodemap_dragmapr_pipeline.R",
  package = "dragmapr"
))
```
